\documentclass[12pt]{letter}
\usepackage{latexsym}

\baselineskip=.15in
\pagestyle{empty}
\signature{Christopher D. Carroll \\ Department of Economics
\\ Johns Hopkins University \\ Baltimore, MD  21218-2685 
\\ 1-410-516-7602 (o) \\ 1-410-516-7600 (f) \\ ccarroll@jhu.edu}

\begin{document}
\begin{letter}{
Philipp Kircher, Editor \\
Review of Economic Studies
}

\opening{Dear Philipp,}

Kjetil Storesletten handled my ``Theoretical Foundations'' paper when
it was originally submitted to ReStud a few years ago (MS 10596); it was
eventually accepted subject to revisions, but in the course of
making those revisions I discovered several ways in which the results
could be made more general, more accessible, and more powerful.  Those
revisions took longer than intended (I was busy, among other things,
working in the Obama administration for a year), but the result is a
paper that is clearly a lot stronger than the draft that was accepted.
However, Kjetil tells me that when he handed his duties over to you,
there was an agreement that any papers that had not met the deadline
for revision (which I have missed by a long margin!)  would be treated
as new submissions.  I understand therefore that the proper procedure
is simply for me to submit the paper through the regular process, as I
am doing?  (I believe I have fully addressed the suggestions of the
referees as well as the editor's suggestions (and in addition have
moved some tedious results to appendices), but as it is not a
resubmission I am not including a formal reply to the referees'
reports or the editor's letter).

At the same time, I would urge you to reconsider a decision that
Kjetil made around time my paper was accepted (subject to revision);
Adam Szeidl had submitted a related paper that fits hand-in-glove with
mine and is quite short, proving the existence of an invariant
distribution in a model of precisely the kind I examine.  (In fact,
Adam's paper was written in response to an earlier draft of my paper
that speculated such a proof was possible but provided only simulation
evidence of invariance).  The two papers are so closely connected that
I would urge you to reconsider Kjetil's rejection of Szeidl's paper;
if you were to do so, I would be able to cut a bit of the discussion
in the last section of my paper (presenting simulation results) by
referring more to Szeidl's paper.

\closing{Sincerely,}

%\ps{}

\end{letter}
\end{document}


